---
title: "Writing Reports with R Markdown"
author: "matthewbrown"
date: "2022-10-21"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(here)
library(tidyverse)
library(knitr)
```

This report was prepared for the UN. It analyzes the relationship between a country's GDP, life expectancy, and CO2 emissions. Our goal was to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that life expectancy and CO2 emissions will increase with a country's GDP.

```{r data, include=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```


```{r gdp_lifeexp_1997, echo=FALSE}
ggplot(gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent,
                           size = pop/1000000)) +
  geom_point() +
  labs(x = "GDP (per capita)", 
                y = "Life Expectancy",
                title = "Do people in wealthy countries live longer?",
                size = "Population (in millions)", 
                color = "Continent")
```

```{r summary_data}
n_countries <- gapminder_1997 %>%
  select(country) %>%
  n_distinct()

min_gdp <- gapminder_1997 %>%
  summarize(min = round(min(gdpPercap)))

max_gdp <- gapminder_1997 %>%
  summarize(max = signif(max(gdpPercap), 1))

```

The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For these countries economic wealth ranged from a minimum of $`r min_gdp` to a maximum of $`r max_gdp` per capita

|Summary of Data|
|--------------------|-----------------|
|Number of countries | `r n_countries` | 
|Minimum GDP per cap | `r min_gdp`     |
|Maximum GDP per cap | `r max_gdp`     |

```{r tables}
gapminder_1997 %>%
  filter(continent == "Oceania") %>%
  kable()


```



## lists 

### unordered lists
* **Python**
  * I don't know Python
* *filter*
* ***automating plot making***
* [Markdown Guide](https://www.markdownguide.org/basic-syntax/)

### ordered lists
1. **Python**
2. *filter*
3. ***automating plot making***
4. Markdown Guide (https://www.markdownguide.org/basic-syntax/)


